LaMachine: A meta-distribution for NLP software

### 2019-01-16 docs/lamachine_clarin2018.tex
work on paper

\documentclass[a4paper,11pt]{article}
\usepackage{etendu-cis}
%\usepackage{CLARIN2018}
% - - - - - - - IMPORTANT - - - - - - - -
% The next three lines allow XeLaTeX, graphics import and hyperlinks, set font and language
%\usepackage{xltxtra,polyglossia,graphicx,hyperref}
%\setmainfont[Mapping=tex-text]{Times}
%\setdefaultlanguage{english}
% If for some reason the above three lines are not compatible with your LaTeX installation,
% comment out the above three instructions and uncomment the following four instead:
\usepackage{times}
\usepackage{hyperref}
\usepackage{graphicx}
%\usepackage{latexsym}
\usepackage[english]{babel}

@@

Once LaMachine is installed, in any of its flavours, you can update it from inside by running
\textttt{lamachine-update}. This updates all of the software managed by LaMachine.

@@

\begin{itemize}
 \item by the Language Machines Research Group and the Centre of Language and Speech Technology, Radboud University Nijmegen
\footnote{Links: \url{https://languagemachines.github.io/timbl}, \url{https://languagemachines.github.io/mbt},
    \url{https://languagemachines.github.io/ucto}, \url{https://languagemachines.github.io/frog},
    \url{http://ilk.uvt.nl/wopr}, \url{https://proycon.github.io/folia}, \url{https://proycon.github.io/clam},
\url{https://github.com/proycon/flat}, \url{https://proycon.anaproy.nl/pynlpl},
\url{https://proycon.github.io/colibri-core/},
\url{https://github.com/proycon/gecco},
\url{https://github.com/proycon/valkuil-gecco},
\url{https://github.com/LanguageMachines/PICCL}, \url{https://githu.com/LanguageMachines/ticcltools},
\url{https://github.com/proycon/labirinto},
\url{https://github.com/proycon/oersetter-webservice}
}
 \begin{itemize}
     \item \textbf{Timbl} - A memory-based machine learning
         toolkit, and \textbf{Mbt}, a memory-based tagger based on timbl.
     \item \textbf{Ucto} - A tokeniser
     \item \textbf{Frog} - Frog is an integration of various memory-based natural language processing (NLP) modules
         developed for Dutch. It can do Part-of-Speech tagging, lemmatisation, named entity recogniton, shallow parsing,
         dependency parsing and morphological analysis. Also included in LaMachine;
         \textbf{Toad}, Trainer Of All Data, training tools for Frog.
     \item \textbf{Wopr}- Memory-based Word Predictor.
     \item \textbf{CLAM} - Quickly build RESTful webservices, powers many webservices offered by LaMachine.
     \item \textbf{FoLiA} - Format for Linguistic Annotation, with tools and libraries in/for Python and $C++$.
     \item \textbf{FLAT} - FoLiA Linguistic Annotation Tool: a web-based linguistic annotation tool.
     \item \textbf{PyNLPl} - Python Natural Language Processing Library.
     \item \textbf{Colibri Core} -- Colibri core is an NLP tool as well as a C++ and Python library for working with
         basic linguistic constructions such as n-grams and skipgrams (i.e patterns with one or more gaps, either of
         fixed or dynamic size) in a quick and memory-efficient way. -
     \item \textbf{Gecco} - Generic Environment for Context-Aware Correction of Orthography - and \textbf{Valkuil.net},
         an instantiation thereof for Dutch.
     \item \textbf{PICCL} - PICCL: A set of workflows for corpus building through OCR, post-correction (through \textbf{TICCL}) and Natural Language Processing.
     \item \textbf{Labirinto}- A web-based portal listing all
         available tools in LaMachine, an ideal starting point for LaMachine.
     \item \textbf{Oersetter} - A Frisian-Dutch Machine Translation system in collaboration with the Fryske Akademy.
 \end{itemize}
 \item by the University of Groningen
 \begin{itemize}
     \item \textbf{Alpino}\footnote{\url{http://www.let.rug.nl/vannoord/alp/Alpino/}} - a dependency parser and tagger for Dutch
 \end{itemize}
 \item by the Vrije Universiteit Amsterdam
 \begin{itemize}
     \item \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} - A python module to parse NAF files
 \end{itemize}
 \item by Utrecht University
 \begin{itemize}
     \item  \textbf{T-scan}\footnote{\url{https://github.com/proycon/tscan}} - T-scan is a Dutch text analytics tool for readability prediction (initially developed at TiCC, Tilburg University, and in collaboration with Radboud University, Nijmegen)
 \end{itemize}
 \item by Meertens Instituut
 \begin{itemize}
     \item \textbf{Python Course for the Humanities}\footnote{http://www.karsdorp.io/python-course/} - Interactive
         tutorial and introduction into programming with Python for the humanities
 \end{itemize}
\end{itemize}

@@

Leveraging this metadata, LaMachine comes with a webserver that offers a portal website with
access and overview of all installed tools, including web services and web applications. It also comes with a
Jupyter Lab\footnote{\url{https://jupyter.org/}} environment which provides a web-based Integrated Development
Environment (IDE) for scripting in Python and R, web-based terminal access, and so-called \emph{notebooks} which mix
text, code and data output and have gained great popularity in data science community nowadays.

>>>>

\documentclass[a4paper,11pt]{article}
\usepackage{CLARIN2018}
% - - - - - - - IMPORTANT - - - - - - - -
% The next three lines allow XeLaTeX, graphics import and hyperlinks, set font and language
%\usepackage{xltxtra,polyglossia,graphicx,hyperref}
%\setmainfont[Mapping=tex-text]{Times}
%\setdefaultlanguage{english}
% If for some reason the above three lines are not compatible with your LaTeX installation,
% comment out the above three instructions and uncomment the following four instead:
\usepackage{times}
\usepackage{hyperref}
\usepackage{graphicx}
%\usepackage{latexsym}
\usepackage[english]{babel}

@@

Once LaMachine is installed, in any of its flavours, you can update it from inside by running
\texttt{lamachine-update}. This updates all of the software managed by LaMachine.

@@

\begin{itemize}
 \item by the Language Machines Research Group and the Centre of Language and Speech Technology, Radboud University
     Nijmegen\footnote{Links: \url{https://languagemachines.github.io/timbl}, \url{https://languagemachines.github.io/mbt},
    \url{https://languagemachines.github.io/ucto}, \url{https://languagemachines.github.io/frog},
    \url{http://ilk.uvt.nl/wopr}, \url{https://proycon.github.io/folia}, \url{https://proycon.github.io/clam},
\url{https://github.com/proycon/flat}, \url{https://proycon.anaproy.nl/pynlpl},
\url{https://proycon.github.io/colibri-core/},
\url{https://github.com/proycon/gecco},
\url{https://github.com/proycon/valkuil-gecco},
\url{https://github.com/LanguageMachines/PICCL}, \url{https://githu.com/LanguageMachines/ticcltools},
\url{https://github.com/proycon/labirinto},
\url{https://github.com/proycon/oersetter-webservice}
}
 \begin{itemize}
     \item \textbf{Timbl} - A memory-based machine learning
         toolkit, and \textbf{Mbt}, a memory-based tagger based on timbl.
     \item \textbf{Ucto} - A tokeniser
     \item \textbf{Frog} - Frog is an integration of various memory-based natural language processing (NLP) modules
         developed for Dutch. It can do Part-of-Speech tagging, lemmatisation, named entity recogniton, shallow parsing,
         dependency parsing and morphological analysis. Also included in LaMachine;
         \textbf{Toad}, Trainer Of All Data, training tools for Frog.
     \item \textbf{Wopr}- Memory-based Word Predictor.
     \item \textbf{CLAM} - Quickly build RESTful webservices, powers many webservices offered by LaMachine.
     \item \textbf{FoLiA} - Format for Linguistic Annotation, with tools and libraries in/for Python and $C++$.
     \item \textbf{FLAT} - FoLiA Linguistic Annotation Tool: a web-based linguistic annotation tool.
     \item \textbf{PyNLPl} - Python Natural Language Processing Library.
     \item \textbf{Colibri Core} -- Colibri core is an NLP tool as well as a C++ and Python library for working with
         basic linguistic constructions such as n-grams and skipgrams (i.e patterns with one or more gaps, either of
         fixed or dynamic size) in a quick and memory-efficient way. -
     \item \textbf{Gecco} - Generic Environment for Context-Aware Correction of Orthography - and \textbf{Valkuil.net},
         an instantiation thereof for Dutch.
     \item \textbf{PICCL} - PICCL: A set of workflows for corpus building through OCR, post-correction (through \textbf{TICCL}) and Natural Language Processing.
     \item \textbf{Labirinto}- A web-based portal listing all
         available tools in LaMachine, an ideal starting point for LaMachine.
     \item \textbf{Oersetter} - A Frisian-Dutch Machine Translation system in collaboration with the Fryske Akademy.
 \end{itemize}
 \item by the University of Groningen
 \begin{itemize}
     \item \textbf{Alpino}\footnote{\url{http://www.let.rug.nl/vannoord/alp/Alpino/}} - a dependency parser and tagger for Dutch
 \end{itemize}
 \item by the Vrije Universiteit Amsterdam
 \begin{itemize}
     \item \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} - A python module to parse NAF files
 \end{itemize}
 \item by Utrecht University
 \begin{itemize}
     \item  \textbf{T-scan}\footnote{\url{https://github.com/proycon/tscan}} - T-scan is a Dutch text analytics tool for readability prediction (initially developed at TiCC, Tilburg University, and in collaboration with Radboud University, Nijmegen)
 \end{itemize}
 \item by Meertens Instituut
 \begin{itemize}
     \item \textbf{Python Course for the Humanities}\footnote{\url{http://www.karsdorp.io/python-course/}} - Interactive
         tutorial and introduction into programming with Python for the humanities
 \end{itemize}
\end{itemize}

@@

LaMachine comes with a webserver that offers a portal website which provides an overview of all installed software in
LaMachine, and acts as a point of access for web services andweb applications. This portal, as shown in
Figure~\ref{fig:portal}, leverages the metadata
registry compiled in each LaMachine installation. In situations where web interfaces are not needed or desired, the
user has the ability to opt-out.

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{screenshot_portal.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the LaMachine portal, powered by Labirinto:
\url{https://github.com/proycon/labirinto}}}
\label{fig:portal}
\end{figure}

LaMachine also comes with  a
Jupyter Lab\footnote{\url{https://jupyter.org/}} environment which provides a web-based Integrated Development
Environment (IDE) for scripting in Python and R, web-based terminal access, and so-called \emph{notebooks} which mix
text, code and data output and have gained great popularity in data science community nowadays. This is shown in
Figure~\ref{fig:lab}.

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{screenshot_lab.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the Jupyter Lab Environment}}
\label{fig:lab}
\end{figure}
